---
title: "week_5_code_along"
author: "Leonardo Feitosa"
date: "09/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(effsize)
library(car)
library(ggbeeswarm)
library(broom)
```

# Part 1. T tests

Mass comparison: male and female chinstrap penguins

```{r}
chinstrap <- penguins %>% 
  filter(species == "Chinstrap")

# Histograms

ggplot(chinstrap, aes(x = body_mass_g)) +
  geom_histogram(bins = 10) +
  facet_wrap(~ sex)

## QQ plot to check out normality
ggplot(chinstrap, aes(sample = body_mass_g)) +
  geom_qq() +
  facet_wrap(~ sex)

## Descriptive statistics
chinstrap %>% 
  group_by(sex) %>% 
  summarise(mean_mass = mean(body_mass_g),
            median_mass = median(body_mass_g),
            sd_mass = sd(body_mass_g),
            sample_size = n())

```


Get vectors of body mass for female and male chinstrap penguins separately

```{r}
chinstraps_f <- chinstrap %>% 
  filter(sex == "female") %>% 
  pull(body_mass_g)


chinstraps_m <- chinstrap %>% 
  filter(sex == "male") %>% 
  pull(body_mass_g)
```

## T-tests a few different ways 

1. Is there significant difference in body mass between male and female chinstrap penguis?

```{r}
chinstrap_ttest <- t.test(chinstraps_f, chinstraps_m)

chinstrap_ttest

cohen.d(chinstraps_f, chinstraps_m)


```


2. Are male chinstrap penguins significantly larger (body mass) than females?

- Order matters here.

```{r}
t.test(chinstraps_m, chinstraps_f, alternative = "greater")
```


## One-way ANOVA

### 1. Table of counts by species and sex

```{r}
penguin_counts <- penguins %>% 
  count(species, sex)
```

Create a subset of only female penguins

```{r}
penguin_f <- penguins %>% 
  filter(sex == "female")


ggplot(penguin_f, aes(x = flipper_length_mm)) +
  geom_histogram(aes(fill = species)) +
  facet_wrap(~ species, scales = "free")


ggplot(penguin_f, aes(sample = flipper_length_mm)) +
  geom_qq(aes(color = species)) +
  facet_wrap(~ species, scales = "free")
```


### Summary statistics

Starting from penguins_f, create a summary table that groups by species and returns the following:

- mean of flipper length
- sd of flipper length
- sample size
- variance `var`
- standard error = sd/sqrt(n)

```{r}
penguin_f_summary <- penguin_f %>%  
  group_by(species) %>% 
  summarise(mean_flipper_length = mean(flipper_length_mm),
            sd_flipper_length = sd(flipper_length_mm),
            sample_size = n(),
            se_flipper_length = sd(flipper_length_mm)/sqrt(n()),
            var_flipper_length = var(flipper_length_mm)
            )


# Levene test
leveneTest(flipper_length_mm ~ species, data = penguin_f)

```


### Beeswarm plot with mean +- 1sd

```{r}
ggplot(penguin_f) +
  geom_beeswarm(aes(x = species, y = flipper_length_mm),
                color = "gray70",
                size = 3,
                alpha = 0.5) +
  geom_point(data = penguin_f_summary,
             aes(x = species, y = mean_flipper_length),
             size = 4) +
  geom_errorbar(data = penguin_f_summary,
                aes(x = species,
                    ymax = mean_flipper_length + sd_flipper_length,
                    ymin = mean_flipper_length - sd_flipper_length),
                width = 0.1) +
  theme_light()
```


### One-way ANOVA

```{r}
penguin_aov <- aov(flipper_length_mm ~ species, data = penguin_f)

summary(penguin_aov)

# Post-hoc test - Tukey's HSD

TukeyHSD(penguin_aov)
```


Type III ANOVA with car package:

```{r}
penguin_lm <- lm(flipper_length_mm ~ species, data = penguin_f)
penguin_aov_t3 <- Anova(penguin_lm, type = "III")
penguin_aov_t3
```










































